---
title: "PM Lung function ACCESS and PROGRESS analysis "
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---





## Load packages and dataset 

```{r, warning=FALSE,message=FALSE}

library(mgcv)
library(nlme)
library(splines2)
library(dlnm)
library (data.table)
library (ggplot2)
library(oddsratio)
library (geepack)
library(doBy)
library (tidyr)
library(psych)
library(plyr)
library(lubridate)

library(jtools)
library(ggiraphExtra)
library(sjPlot)
library(sjmisc)
library(interplot)
library(TMB)

library (knitr)
library(kableExtra)
library (ggpubr)


```

## Separate cohorts 


## PROGRESS


```{r, message=FALSE, warning=FALSE}


library(haven)
PROGRESS_ACCESS_Lung_function_PM_Temp_ <- read_sav("PROGRESS ACCESS PM temp month avgs 020722 small_3.sav")

main<-subset(PROGRESS_ACCESS_Lung_function_PM_Temp_,  cohort_num==1)

attach(main)

```


## Create penalized crossbasis 

```{r}

pmweek<-as.matrix (main[,5242:5315])

                         
pm_exp_cb <- crossbasis(pmweek, 
                          lag = c(1, 74), 
                          argvar = list(fun = 'lin'),
                          arglag = list(fun = 'cr'))




tempweek<-as.matrix (main[,5328:5401])

                         
temp_exp_cb <- crossbasis(tempweek, 
                          lag = c(1, 74), 
                          argvar = list(fun = 'lin'),
                          arglag = list(fun = 'cr'))
                          
pen1 <- cbPen(pm_exp_cb )
pen2 <- cbPen(temp_exp_cb)


```





## Method=REML

## FEV model 


```{r, echo=TRUE, warning=FALSE}





mod  <- gam(z_pre_fev1 ~pm_exp_cb + + temp_exp_cb   +as.factor (ets_postnatal) +mom_age +as.factor (educ_dich)   ,paraPen = list(pm_exp_cb  = pen1,  
                          temp_exp_cb  = pen2), 
           method = 'REML') # Estimation technique
summary (mod)

# Summarize results
pre_exp_cr <- crossreduce(pm_exp_cb , # DLM crossbasis used in model
                          mod, # Model results
                          type = "var", # Reduction across exposure values
                          value = 5, # Effect relative to exposure value = 1
                          ci.level = 0.95) # Confidence level
post_exp_cr <- crossreduce(temp_exp_cb , mod, type = "var", value = 5)
pre_dlm_results <- data.frame("fit" = pre_exp_cr$fit, 
                              "low" = pre_exp_cr$low,
                              "high" = pre_exp_cr$high)
post_dlm_results <- data.frame("fit" = post_exp_cr$fit, 
                              "low" = post_exp_cr$low,
                              "high" = post_exp_cr$high)



# name rows with '*' to specify critical windows
row.names(pre_dlm_results) <- ifelse(pre_dlm_results$low > 0 | pre_dlm_results$high < 0, paste0("*", row.names(pre_dlm_results)), row.names(pre_dlm_results))
row.names(post_dlm_results) <- ifelse(post_dlm_results$low > 0 | post_dlm_results$high < 0, paste0("*", row.names(post_dlm_results)), row.names(post_dlm_results))
```



```{r, warning=FALSE,message=FALSE}
round(pre_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```

```{r, warning=FALSE,message=FALSE}
round(post_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```




```{r, warning=FALSE,message=FALSE}
pre_dlm_results$lag = 1:74
p1 <- ggplot(pre_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEV1 z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 
 

post_dlm_results$lag = 1:74
p2 <- ggplot(post_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEV1 z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 

pFEV1_ageEdu = ggarrange(p1, p2)
pFEV1_ageEdu

p1

p2



```






## Method=REML

## FVC model 


```{r, echo=TRUE, warning=FALSE}





mod  <- gam(z_pre_fvc ~pm_exp_cb + + temp_exp_cb   +as.factor (ets_postnatal) +mom_age +as.factor (educ_dich)  ,paraPen = list(pm_exp_cb  = pen1,  
                          temp_exp_cb  = pen2), 
           method = 'REML') # Estimation technique

summary (mod)


# Summarize results
pre_exp_cr <- crossreduce(pm_exp_cb , # DLM crossbasis used in model
                          mod, # Model results
                          type = "var", # Reduction across exposure values
                          value = 5, # Effect relative to exposure value = 1
                          ci.level = 0.95) # Confidence level
post_exp_cr <- crossreduce(temp_exp_cb , mod, type = "var", value = 5)
pre_dlm_results <- data.frame("fit" = pre_exp_cr$fit, 
                              "low" = pre_exp_cr$low,
                              "high" = pre_exp_cr$high)
post_dlm_results <- data.frame("fit" = post_exp_cr$fit, 
                              "low" = post_exp_cr$low,
                              "high" = post_exp_cr$high)
# name rows with '*' to specify critical windows
row.names(pre_dlm_results) <- ifelse(pre_dlm_results$low > 0 | pre_dlm_results$high < 0, paste0("*", row.names(pre_dlm_results)), row.names(pre_dlm_results))
row.names(post_dlm_results) <- ifelse(post_dlm_results$low > 0 | post_dlm_results$high < 0, paste0("*", row.names(post_dlm_results)), row.names(post_dlm_results))
```


```{r, warning=FALSE,message=FALSE}
round(pre_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```

```{r, warning=FALSE,message=FALSE}
round(post_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```




```{r, warning=FALSE,message=FALSE}
pre_dlm_results$lag = 1:74
p1 <- ggplot(pre_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FVC z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 
 

post_dlm_results$lag = 1:74
p2 <- ggplot(post_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FVC z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 

pFEV1_ageEdu = ggarrange(p1, p2)
pFEV1_ageEdu

p1
p2


```




## Method=REML

## FEV1/FVC model 


```{r, echo=TRUE, warning=FALSE}





mod  <- gam(z_pre_fev1_fvc2 ~pm_exp_cb + + temp_exp_cb   +as.factor (ets_postnatal) +mom_age +as.factor (educ_dich)  ,paraPen = list(pm_exp_cb  = pen1,  
                          temp_exp_cb  = pen2), 
           method = 'REML') # Estimation technique

summary (mod)


# Summarize results
pre_exp_cr <- crossreduce(pm_exp_cb , # DLM crossbasis used in model
                          mod, # Model results
                          type = "var", # Reduction across exposure values
                          value = 5, # Effect relative to exposure value = 1
                          ci.level = 0.95) # Confidence level
post_exp_cr <- crossreduce(temp_exp_cb , mod, type = "var", value = 5)
pre_dlm_results <- data.frame("fit" = pre_exp_cr$fit, 
                              "low" = pre_exp_cr$low,
                              "high" = pre_exp_cr$high)
post_dlm_results <- data.frame("fit" = post_exp_cr$fit, 
                              "low" = post_exp_cr$low,
                              "high" = post_exp_cr$high)
# name rows with '*' to specify critical windows
row.names(pre_dlm_results) <- ifelse(pre_dlm_results$low > 0 | pre_dlm_results$high < 0, paste0("*", row.names(pre_dlm_results)), row.names(pre_dlm_results))
row.names(post_dlm_results) <- ifelse(post_dlm_results$low > 0 | post_dlm_results$high < 0, paste0("*", row.names(post_dlm_results)), row.names(post_dlm_results))
```


```{r, warning=FALSE,message=FALSE}
round(pre_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```

```{r, warning=FALSE,message=FALSE}
round(post_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```




```{r, warning=FALSE,message=FALSE}
pre_dlm_results$lag = 1:74
p1 <- ggplot(pre_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEV1/FVC z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 
 

post_dlm_results$lag = 1:74
p2 <- ggplot(post_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEV1/FVC z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 

pFEV1_ageEdu = ggarrange(p1, p2)
pFEV1_ageEdu

p1
p2


```





## Method=REML

## FEF model 


```{r, echo=TRUE, warning=FALSE}





mod  <- gam(z_pre_fef ~pm_exp_cb + + temp_exp_cb   +as.factor (ets_postnatal) +mom_age +as.factor (educ_dich)   ,paraPen = list(pm_exp_cb  = pen1,  
                          temp_exp_cb  = pen2), 
           method = 'REML') # Estimation technique

summary (mod)


# Summarize results
pre_exp_cr <- crossreduce(pm_exp_cb , # DLM crossbasis used in model
                          mod, # Model results
                          type = "var", # Reduction across exposure values
                          value = 5, # Effect relative to exposure value = 1
                          ci.level = 0.95) # Confidence level
post_exp_cr <- crossreduce(temp_exp_cb , mod, type = "var", value = 5)
pre_dlm_results <- data.frame("fit" = pre_exp_cr$fit, 
                              "low" = pre_exp_cr$low,
                              "high" = pre_exp_cr$high)
post_dlm_results <- data.frame("fit" = post_exp_cr$fit, 
                              "low" = post_exp_cr$low,
                              "high" = post_exp_cr$high)
# name rows with '*' to specify critical windows
row.names(pre_dlm_results) <- ifelse(pre_dlm_results$low > 0 | pre_dlm_results$high < 0, paste0("*", row.names(pre_dlm_results)), row.names(pre_dlm_results))
row.names(post_dlm_results) <- ifelse(post_dlm_results$low > 0 | post_dlm_results$high < 0, paste0("*", row.names(post_dlm_results)), row.names(post_dlm_results))
```


```{r, warning=FALSE,message=FALSE}
round(pre_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```

```{r, warning=FALSE,message=FALSE}
round(post_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```




```{r, warning=FALSE,message=FALSE}
pre_dlm_results$lag = 1:74
p1 <- ggplot(pre_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEF z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 
 

post_dlm_results$lag = 1:74
p2 <- ggplot(post_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEF z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 

pFEV1_ageEdu = ggarrange(p1, p2)
pFEV1_ageEdu

p1
p2



```




## ACCESS

```{r, message=FALSE, warning=FALSE}


library(haven)
PROGRESS_ACCESS_Lung_function_PM_Temp_ <- read_sav("PROGRESS ACCESS PM temp month avgs 020722 small_3.sav")

main<-subset(PROGRESS_ACCESS_Lung_function_PM_Temp_,  cohort_num==0)

attach(main)
```



## Create penalized crossbasis 

```{r}

pmweek<-as.matrix (main[,5242:5315])

                         
pm_exp_cb <- crossbasis(pmweek, 
                          lag = c(1, 74), 
                          argvar = list(fun = 'lin'),
                          arglag = list(fun = 'cr'))




tempweek<-as.matrix (main[,5328:5401])

                         
temp_exp_cb <- crossbasis(tempweek, 
                          lag = c(1, 74), 
                          argvar = list(fun = 'lin'),
                          arglag = list(fun = 'cr'))
                          
pen1 <- cbPen(pm_exp_cb )
pen2 <- cbPen(temp_exp_cb)


```





## Method=REML

## FEV model 


```{r, echo=TRUE, warning=FALSE}





mod  <- gam(z_pre_fev1 ~pm_exp_cb + + temp_exp_cb   +as.factor (ets_postnatal) +mom_age +as.factor (educ_dich)   ,paraPen = list(pm_exp_cb  = pen1,  
                          temp_exp_cb  = pen2), 
           method = 'REML') # Estimation technique
summary (mod)

# Summarize results
pre_exp_cr <- crossreduce(pm_exp_cb , # DLM crossbasis used in model
                          mod, # Model results
                          type = "var", # Reduction across exposure values
                          value = 5, # Effect relative to exposure value = 1
                          ci.level = 0.95) # Confidence level
post_exp_cr <- crossreduce(temp_exp_cb , mod, type = "var", value = 5)
pre_dlm_results <- data.frame("fit" = pre_exp_cr$fit, 
                              "low" = pre_exp_cr$low,
                              "high" = pre_exp_cr$high)
post_dlm_results <- data.frame("fit" = post_exp_cr$fit, 
                              "low" = post_exp_cr$low,
                              "high" = post_exp_cr$high)



# name rows with '*' to specify critical windows
row.names(pre_dlm_results) <- ifelse(pre_dlm_results$low > 0 | pre_dlm_results$high < 0, paste0("*", row.names(pre_dlm_results)), row.names(pre_dlm_results))
row.names(post_dlm_results) <- ifelse(post_dlm_results$low > 0 | post_dlm_results$high < 0, paste0("*", row.names(post_dlm_results)), row.names(post_dlm_results))
```


```{r, warning=FALSE,message=FALSE}
round(pre_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```

```{r, warning=FALSE,message=FALSE}
round(post_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```




```{r, warning=FALSE,message=FALSE}
pre_dlm_results$lag = 1:74
p1 <- ggplot(pre_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEV1 z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 
 

post_dlm_results$lag = 1:74
p2 <- ggplot(post_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEV1 z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 

pFEV1_ageEdu = ggarrange(p1, p2)
pFEV1_ageEdu

p1

p2



```






## Method=REML

## FVC model 


```{r, echo=TRUE, warning=FALSE}





mod  <- gam(z_pre_fvc ~pm_exp_cb + + temp_exp_cb   +as.factor (ets_postnatal) +mom_age +as.factor (educ_dich)  ,paraPen = list(pm_exp_cb  = pen1,  
                          temp_exp_cb  = pen2), 
           method = 'REML') # Estimation technique

summary (mod)


# Summarize results
pre_exp_cr <- crossreduce(pm_exp_cb , # DLM crossbasis used in model
                          mod, # Model results
                          type = "var", # Reduction across exposure values
                          value = 5, # Effect relative to exposure value = 1
                          ci.level = 0.95) # Confidence level
post_exp_cr <- crossreduce(temp_exp_cb , mod, type = "var", value = 5)
pre_dlm_results <- data.frame("fit" = pre_exp_cr$fit, 
                              "low" = pre_exp_cr$low,
                              "high" = pre_exp_cr$high)
post_dlm_results <- data.frame("fit" = post_exp_cr$fit, 
                              "low" = post_exp_cr$low,
                              "high" = post_exp_cr$high)
# name rows with '*' to specify critical windows
row.names(pre_dlm_results) <- ifelse(pre_dlm_results$low > 0 | pre_dlm_results$high < 0, paste0("*", row.names(pre_dlm_results)), row.names(pre_dlm_results))
row.names(post_dlm_results) <- ifelse(post_dlm_results$low > 0 | post_dlm_results$high < 0, paste0("*", row.names(post_dlm_results)), row.names(post_dlm_results))
```


```{r, warning=FALSE,message=FALSE}
round(pre_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```

```{r, warning=FALSE,message=FALSE}
round(post_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```




```{r, warning=FALSE,message=FALSE}
pre_dlm_results$lag = 1:74
p1 <- ggplot(pre_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FVC z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 
 

post_dlm_results$lag = 1:74
p2 <- ggplot(post_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FVC z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 

pFEV1_ageEdu = ggarrange(p1, p2)
pFEV1_ageEdu

p1
p2


```




## Method=REML

## FEV1/FVC model 


```{r, echo=TRUE, warning=FALSE}





mod  <- gam(z_pre_fev1_fvc2 ~pm_exp_cb + + temp_exp_cb   +as.factor (ets_postnatal) +mom_age +as.factor (educ_dich)  ,paraPen = list(pm_exp_cb  = pen1,  
                          temp_exp_cb  = pen2), 
           method = 'REML') # Estimation technique

summary (mod)


# Summarize results
pre_exp_cr <- crossreduce(pm_exp_cb , # DLM crossbasis used in model
                          mod, # Model results
                          type = "var", # Reduction across exposure values
                          value = 5, # Effect relative to exposure value = 1
                          ci.level = 0.95) # Confidence level
post_exp_cr <- crossreduce(temp_exp_cb , mod, type = "var", value = 5)
pre_dlm_results <- data.frame("fit" = pre_exp_cr$fit, 
                              "low" = pre_exp_cr$low,
                              "high" = pre_exp_cr$high)
post_dlm_results <- data.frame("fit" = post_exp_cr$fit, 
                              "low" = post_exp_cr$low,
                              "high" = post_exp_cr$high)
# name rows with '*' to specify critical windows
row.names(pre_dlm_results) <- ifelse(pre_dlm_results$low > 0 | pre_dlm_results$high < 0, paste0("*", row.names(pre_dlm_results)), row.names(pre_dlm_results))
row.names(post_dlm_results) <- ifelse(post_dlm_results$low > 0 | post_dlm_results$high < 0, paste0("*", row.names(post_dlm_results)), row.names(post_dlm_results))
```


```{r, warning=FALSE,message=FALSE}
round(pre_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```

```{r, warning=FALSE,message=FALSE}
round(post_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```




```{r, warning=FALSE,message=FALSE}
pre_dlm_results$lag = 1:74
p1 <- ggplot(pre_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEV1/FVC z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 
 

post_dlm_results$lag = 1:74
p2 <- ggplot(post_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEV1/FVC z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 

pFEV1_ageEdu = ggarrange(p1, p2)
pFEV1_ageEdu

p1
p2


```





## Method=REML

## FEF model 


```{r, echo=TRUE, warning=FALSE}





mod  <- gam(z_pre_fef ~pm_exp_cb + + temp_exp_cb   +as.factor (ets_postnatal) +mom_age +as.factor (educ_dich)   ,paraPen = list(pm_exp_cb  = pen1,  
                          temp_exp_cb  = pen2), 
           method = 'REML') # Estimation technique

summary (mod)


# Summarize results
pre_exp_cr <- crossreduce(pm_exp_cb , # DLM crossbasis used in model
                          mod, # Model results
                          type = "var", # Reduction across exposure values
                          value = 5, # Effect relative to exposure value = 1
                          ci.level = 0.95) # Confidence level
post_exp_cr <- crossreduce(temp_exp_cb , mod, type = "var", value = 1)
pre_dlm_results <- data.frame("fit" = pre_exp_cr$fit, 
                              "low" = pre_exp_cr$low,
                              "high" = pre_exp_cr$high)
post_dlm_results <- data.frame("fit" = post_exp_cr$fit, 
                              "low" = post_exp_cr$low,
                              "high" = post_exp_cr$high)
# name rows with '*' to specify critical windows
row.names(pre_dlm_results) <- ifelse(pre_dlm_results$low > 0 | pre_dlm_results$high < 0, paste0("*", row.names(pre_dlm_results)), row.names(pre_dlm_results))
row.names(post_dlm_results) <- ifelse(post_dlm_results$low > 0 | post_dlm_results$high < 0, paste0("*", row.names(post_dlm_results)), row.names(post_dlm_results))
```


```{r, warning=FALSE,message=FALSE}
round(pre_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```

```{r, warning=FALSE,message=FALSE}
round(post_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```




```{r, warning=FALSE,message=FALSE}
pre_dlm_results$lag = 1:74
p1 <- ggplot(pre_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEF z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 
 

post_dlm_results$lag = 1:74
p2 <- ggplot(post_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEF z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 

pFEV1_ageEdu = ggarrange(p1, p2)
pFEV1_ageEdu

p1
p2



```


## combined cohorts with only adjustment for cohort variable


```{r, message=FALSE, warning=FALSE}


library(haven)
PROGRESS_ACCESS_Lung_function_PM_Temp_ <- read_sav("PROGRESS ACCESS PM temp month avgs 020722 small_3.sav")

main<-subset(PROGRESS_ACCESS_Lung_function_PM_Temp_)

attach(main)

```

## Create penalized crossbasis 

```{r}

pmweek<-as.matrix (main[,5242:5315])

                         
pm_exp_cb <- crossbasis(pmweek, 
                          lag = c(1, 74), 
                          argvar = list(fun = 'lin'),
                          arglag = list(fun = 'cr'))




tempweek<-as.matrix (main[,5328:5401])

                         
temp_exp_cb <- crossbasis(tempweek, 
                          lag = c(1, 74), 
                          argvar = list(fun = 'lin'),
                          arglag = list(fun = 'cr'))
                          
pen1 <- cbPen(pm_exp_cb )
pen2 <- cbPen(temp_exp_cb)


```





## Method=REML

## FEV model 


```{r, echo=TRUE, warning=FALSE}





mod  <- gam(z_pre_fev1 ~pm_exp_cb + + temp_exp_cb   +as.factor (ets_postnatal) +mom_age +as.factor (educ_dich) +as.factor(cohort)   ,paraPen = list(pm_exp_cb  = pen1,  
                          temp_exp_cb  = pen2), 
           method = 'REML') # Estimation technique
summary (mod)

# Summarize results
pre_exp_cr <- crossreduce(pm_exp_cb , # DLM crossbasis used in model
                          mod, # Model results
                          type = "var", # Reduction across exposure values
                          value = 5, # Effect relative to exposure value = 1
                          ci.level = 0.95) # Confidence level
post_exp_cr <- crossreduce(temp_exp_cb , mod, type = "var", value = 5)
pre_dlm_results <- data.frame("fit" = pre_exp_cr$fit, 
                              "low" = pre_exp_cr$low,
                              "high" = pre_exp_cr$high)
post_dlm_results <- data.frame("fit" = post_exp_cr$fit, 
                              "low" = post_exp_cr$low,
                              "high" = post_exp_cr$high)



# name rows with '*' to specify critical windows
row.names(pre_dlm_results) <- ifelse(pre_dlm_results$low > 0 | pre_dlm_results$high < 0, paste0("*", row.names(pre_dlm_results)), row.names(pre_dlm_results))
row.names(post_dlm_results) <- ifelse(post_dlm_results$low > 0 | post_dlm_results$high < 0, paste0("*", row.names(post_dlm_results)), row.names(post_dlm_results))
```


```{r, warning=FALSE,message=FALSE}
round(pre_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```

```{r, warning=FALSE,message=FALSE}
round(post_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```




```{r, warning=FALSE,message=FALSE}
pre_dlm_results$lag = 1:74
p1 <- ggplot(pre_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEV1 z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 
 

post_dlm_results$lag = 1:74
p2 <- ggplot(post_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEV1 z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 

pFEV1_ageEdu = ggarrange(p1, p2)
pFEV1_ageEdu

p1

p2



```






## Method=REML

## FVC model 


```{r, echo=TRUE, warning=FALSE}





mod  <- gam(z_pre_fvc ~pm_exp_cb + + temp_exp_cb   +as.factor (ets_postnatal) +mom_age +as.factor (educ_dich) +as.factor(cohort)  ,paraPen = list(pm_exp_cb  = pen1,  
                          temp_exp_cb  = pen2), 
           method = 'REML') # Estimation technique

summary (mod)


# Summarize results
pre_exp_cr <- crossreduce(pm_exp_cb , # DLM crossbasis used in model
                          mod, # Model results
                          type = "var", # Reduction across exposure values
                          value = 5, # Effect relative to exposure value = 1
                          ci.level = 0.95) # Confidence level
post_exp_cr <- crossreduce(temp_exp_cb , mod, type = "var", value = 5)
pre_dlm_results <- data.frame("fit" = pre_exp_cr$fit, 
                              "low" = pre_exp_cr$low,
                              "high" = pre_exp_cr$high)
post_dlm_results <- data.frame("fit" = post_exp_cr$fit, 
                              "low" = post_exp_cr$low,
                              "high" = post_exp_cr$high)
# name rows with '*' to specify critical windows
row.names(pre_dlm_results) <- ifelse(pre_dlm_results$low > 0 | pre_dlm_results$high < 0, paste0("*", row.names(pre_dlm_results)), row.names(pre_dlm_results))
row.names(post_dlm_results) <- ifelse(post_dlm_results$low > 0 | post_dlm_results$high < 0, paste0("*", row.names(post_dlm_results)), row.names(post_dlm_results))
```


```{r, warning=FALSE,message=FALSE}
round(pre_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```

```{r, warning=FALSE,message=FALSE}
round(post_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```




```{r, warning=FALSE,message=FALSE}
pre_dlm_results$lag = 1:74
p1 <- ggplot(pre_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FVC z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 
 

post_dlm_results$lag = 1:74
p2 <- ggplot(post_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FVC z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 

pFEV1_ageEdu = ggarrange(p1, p2)
pFEV1_ageEdu

p1
p2


```




## Method=REML

## FEV1/FVC model 


```{r, echo=TRUE, warning=FALSE}





mod  <- gam(z_pre_fev1_fvc2 ~pm_exp_cb + + temp_exp_cb   +as.factor (ets_postnatal) +mom_age +as.factor (educ_dich) +as.factor(cohort)  ,paraPen = list(pm_exp_cb  = pen1,  
                          temp_exp_cb  = pen2), 
           method = 'REML') # Estimation technique

summary (mod)


# Summarize results
pre_exp_cr <- crossreduce(pm_exp_cb , # DLM crossbasis used in model
                          mod, # Model results
                          type = "var", # Reduction across exposure values
                          value = 5, # Effect relative to exposure value = 1
                          ci.level = 0.95) # Confidence level
post_exp_cr <- crossreduce(temp_exp_cb , mod, type = "var", value = 5)
pre_dlm_results <- data.frame("fit" = pre_exp_cr$fit, 
                              "low" = pre_exp_cr$low,
                              "high" = pre_exp_cr$high)
post_dlm_results <- data.frame("fit" = post_exp_cr$fit, 
                              "low" = post_exp_cr$low,
                              "high" = post_exp_cr$high)
# name rows with '*' to specify critical windows
row.names(pre_dlm_results) <- ifelse(pre_dlm_results$low > 0 | pre_dlm_results$high < 0, paste0("*", row.names(pre_dlm_results)), row.names(pre_dlm_results))
row.names(post_dlm_results) <- ifelse(post_dlm_results$low > 0 | post_dlm_results$high < 0, paste0("*", row.names(post_dlm_results)), row.names(post_dlm_results))
```


```{r, warning=FALSE,message=FALSE}
round(pre_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```

```{r, warning=FALSE,message=FALSE}
round(post_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```




```{r, warning=FALSE,message=FALSE}
pre_dlm_results$lag = 1:74
p1 <- ggplot(pre_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEV1/FVC z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 
 

post_dlm_results$lag = 1:74
p2 <- ggplot(post_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEV1/FVC z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 

pFEV1_ageEdu = ggarrange(p1, p2)
pFEV1_ageEdu

p1
p2


```





## Method=REML

## FEF model 


```{r, echo=TRUE, warning=FALSE}





mod  <- gam(z_pre_fef ~pm_exp_cb + + temp_exp_cb   +as.factor (ets_postnatal) +mom_age +as.factor (educ_dich) +as.factor(cohort)   ,paraPen = list(pm_exp_cb  = pen1,  
                          temp_exp_cb  = pen2), 
           method = 'REML') # Estimation technique

summary (mod)


# Summarize results
pre_exp_cr <- crossreduce(pm_exp_cb , # DLM crossbasis used in model
                          mod, # Model results
                          type = "var", # Reduction across exposure values
                          value = 5, # Effect relative to exposure value = 1
                          ci.level = 0.95) # Confidence level
post_exp_cr <- crossreduce(temp_exp_cb , mod, type = "var", value = 5)
pre_dlm_results <- data.frame("fit" = pre_exp_cr$fit, 
                              "low" = pre_exp_cr$low,
                              "high" = pre_exp_cr$high)
post_dlm_results <- data.frame("fit" = post_exp_cr$fit, 
                              "low" = post_exp_cr$low,
                              "high" = post_exp_cr$high)
# name rows with '*' to specify critical windows
row.names(pre_dlm_results) <- ifelse(pre_dlm_results$low > 0 | pre_dlm_results$high < 0, paste0("*", row.names(pre_dlm_results)), row.names(pre_dlm_results))
row.names(post_dlm_results) <- ifelse(post_dlm_results$low > 0 | post_dlm_results$high < 0, paste0("*", row.names(post_dlm_results)), row.names(post_dlm_results))
```


```{r, warning=FALSE,message=FALSE}
round(pre_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```

```{r, warning=FALSE,message=FALSE}
round(post_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```




```{r, warning=FALSE,message=FALSE}
pre_dlm_results$lag = 1:74
p1 <- ggplot(pre_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEF z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 
 

post_dlm_results$lag = 1:74
p2 <- ggplot(post_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEF z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 

pFEV1_ageEdu = ggarrange(p1, p2)
pFEV1_ageEdu

p1
p2



```




## using Caliper method  

```{r, message=FALSE, warning=FALSE}


library(haven)
PROGRESS_ACCESS_Lung_function_PM_Temp_ <- read_sav("PROGRESS ACCESS PM temp month avgs 020722 small_3.sav")

main<-subset(PROGRESS_ACCESS_Lung_function_PM_Temp_, caliper3>=1)

attach(main)

```


## Create penalized crossbasis 

```{r}

pmweek<-as.matrix (main[,5242:5315])

                         
pm_exp_cb <- crossbasis(pmweek, 
                          lag = c(1, 74), 
                          argvar = list(fun = 'lin'),
                          arglag = list(fun = 'cr'))




tempweek<-as.matrix (main[,5328:5401])

                         
temp_exp_cb <- crossbasis(tempweek, 
                          lag = c(1, 74), 
                          argvar = list(fun = 'lin'),
                          arglag = list(fun = 'cr'))
                          
pen1 <- cbPen(pm_exp_cb )
pen2 <- cbPen(temp_exp_cb)


```





## Method=REML

## FEV model 


```{r, echo=TRUE, warning=FALSE}





mod  <- gam(z_pre_fev1 ~pm_exp_cb + + temp_exp_cb   +as.factor (ets_postnatal) +mom_age +as.factor (educ_dich)   ,paraPen = list(pm_exp_cb  = pen1,  
                          temp_exp_cb  = pen2), 
           method = 'REML') # Estimation technique
summary (mod)

# Summarize results
pre_exp_cr <- crossreduce(pm_exp_cb , # DLM crossbasis used in model
                          mod, # Model results
                          type = "var", # Reduction across exposure values
                          value = 5, # Effect relative to exposure value = 1
                          ci.level = 0.95) # Confidence level
post_exp_cr <- crossreduce(temp_exp_cb , mod, type = "var", value = 5)
pre_dlm_results <- data.frame("fit" = pre_exp_cr$fit, 
                              "low" = pre_exp_cr$low,
                              "high" = pre_exp_cr$high)
post_dlm_results <- data.frame("fit" = post_exp_cr$fit, 
                              "low" = post_exp_cr$low,
                              "high" = post_exp_cr$high)



# name rows with '*' to specify critical windows
row.names(pre_dlm_results) <- ifelse(pre_dlm_results$low > 0 | pre_dlm_results$high < 0, paste0("*", row.names(pre_dlm_results)), row.names(pre_dlm_results))
row.names(post_dlm_results) <- ifelse(post_dlm_results$low > 0 | post_dlm_results$high < 0, paste0("*", row.names(post_dlm_results)), row.names(post_dlm_results))
```



```{r, warning=FALSE,message=FALSE}
round(pre_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```

```{r, warning=FALSE,message=FALSE}
round(post_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```




```{r, warning=FALSE,message=FALSE}
pre_dlm_results$lag = 1:74
p1 <- ggplot(pre_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEV1 z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 
 

post_dlm_results$lag = 1:74
p2 <- ggplot(post_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEV1 z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 

pFEV1_ageEdu = ggarrange(p1, p2)
pFEV1_ageEdu

p1

p2



```






## Method=REML

## FVC model 


```{r, echo=TRUE, warning=FALSE}





mod  <- gam(z_pre_fvc ~pm_exp_cb + + temp_exp_cb   +as.factor (ets_postnatal) +mom_age +as.factor (educ_dich)  ,paraPen = list(pm_exp_cb  = pen1,  
                          temp_exp_cb  = pen2), 
           method = 'REML') # Estimation technique

summary (mod)


# Summarize results
pre_exp_cr <- crossreduce(pm_exp_cb , # DLM crossbasis used in model
                          mod, # Model results
                          type = "var", # Reduction across exposure values
                          value = 5, # Effect relative to exposure value = 1
                          ci.level = 0.95) # Confidence level
post_exp_cr <- crossreduce(temp_exp_cb , mod, type = "var", value = 5)
pre_dlm_results <- data.frame("fit" = pre_exp_cr$fit, 
                              "low" = pre_exp_cr$low,
                              "high" = pre_exp_cr$high)
post_dlm_results <- data.frame("fit" = post_exp_cr$fit, 
                              "low" = post_exp_cr$low,
                              "high" = post_exp_cr$high)
# name rows with '*' to specify critical windows
row.names(pre_dlm_results) <- ifelse(pre_dlm_results$low > 0 | pre_dlm_results$high < 0, paste0("*", row.names(pre_dlm_results)), row.names(pre_dlm_results))
row.names(post_dlm_results) <- ifelse(post_dlm_results$low > 0 | post_dlm_results$high < 0, paste0("*", row.names(post_dlm_results)), row.names(post_dlm_results))
```


```{r, warning=FALSE,message=FALSE}
round(pre_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```

```{r, warning=FALSE,message=FALSE}
round(post_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```




```{r, warning=FALSE,message=FALSE}
pre_dlm_results$lag = 1:74
p1 <- ggplot(pre_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FVC z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 
 

post_dlm_results$lag = 1:74
p2 <- ggplot(post_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FVC z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 

pFEV1_ageEdu = ggarrange(p1, p2)
pFEV1_ageEdu

p1
p2


```




## Method=REML

## FEV1/FVC model 


```{r, echo=TRUE, warning=FALSE}





mod  <- gam(z_pre_fev1_fvc2 ~pm_exp_cb + + temp_exp_cb   +as.factor (ets_postnatal) +mom_age +as.factor (educ_dich)  ,paraPen = list(pm_exp_cb  = pen1,  
                          temp_exp_cb  = pen2), 
           method = 'REML') # Estimation technique

summary (mod)


# Summarize results
pre_exp_cr <- crossreduce(pm_exp_cb , # DLM crossbasis used in model
                          mod, # Model results
                          type = "var", # Reduction across exposure values
                          value = 5, # Effect relative to exposure value = 1
                          ci.level = 0.95) # Confidence level
post_exp_cr <- crossreduce(temp_exp_cb , mod, type = "var", value = 5)
pre_dlm_results <- data.frame("fit" = pre_exp_cr$fit, 
                              "low" = pre_exp_cr$low,
                              "high" = pre_exp_cr$high)
post_dlm_results <- data.frame("fit" = post_exp_cr$fit, 
                              "low" = post_exp_cr$low,
                              "high" = post_exp_cr$high)
# name rows with '*' to specify critical windows
row.names(pre_dlm_results) <- ifelse(pre_dlm_results$low > 0 | pre_dlm_results$high < 0, paste0("*", row.names(pre_dlm_results)), row.names(pre_dlm_results))
row.names(post_dlm_results) <- ifelse(post_dlm_results$low > 0 | post_dlm_results$high < 0, paste0("*", row.names(post_dlm_results)), row.names(post_dlm_results))
```


```{r, warning=FALSE,message=FALSE}
round(pre_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```

```{r, warning=FALSE,message=FALSE}
round(post_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```




```{r, warning=FALSE,message=FALSE}
pre_dlm_results$lag = 1:74
p1 <- ggplot(pre_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEV1/FVC z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 
 

post_dlm_results$lag = 1:74
p2 <- ggplot(post_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEV1/FVC z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 

pFEV1_ageEdu = ggarrange(p1, p2)
pFEV1_ageEdu

p1
p2


```





## Method=REML

## FEF model 


```{r, echo=TRUE, warning=FALSE}





mod  <- gam(z_pre_fef ~pm_exp_cb + + temp_exp_cb   +as.factor (ets_postnatal) +mom_age +as.factor (educ_dich)   ,paraPen = list(pm_exp_cb  = pen1,  
                          temp_exp_cb  = pen2), 
           method = 'REML') # Estimation technique

summary (mod)


# Summarize results
pre_exp_cr <- crossreduce(pm_exp_cb , # DLM crossbasis used in model
                          mod, # Model results
                          type = "var", # Reduction across exposure values
                          value = 5, # Effect relative to exposure value = 1
                          ci.level = 0.95) # Confidence level
post_exp_cr <- crossreduce(temp_exp_cb , mod, type = "var", value = 5)
pre_dlm_results <- data.frame("fit" = pre_exp_cr$fit, 
                              "low" = pre_exp_cr$low,
                              "high" = pre_exp_cr$high)
post_dlm_results <- data.frame("fit" = post_exp_cr$fit, 
                              "low" = post_exp_cr$low,
                              "high" = post_exp_cr$high)
# name rows with '*' to specify critical windows
row.names(pre_dlm_results) <- ifelse(pre_dlm_results$low > 0 | pre_dlm_results$high < 0, paste0("*", row.names(pre_dlm_results)), row.names(pre_dlm_results))
row.names(post_dlm_results) <- ifelse(post_dlm_results$low > 0 | post_dlm_results$high < 0, paste0("*", row.names(post_dlm_results)), row.names(post_dlm_results))
```


```{r, warning=FALSE,message=FALSE}
round(pre_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```

```{r, warning=FALSE,message=FALSE}
round(post_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```




```{r, warning=FALSE,message=FALSE}
pre_dlm_results$lag = 1:74
p1 <- ggplot(pre_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEF z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 
 

post_dlm_results$lag = 1:74
p2 <- ggplot(post_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEF z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 

pFEV1_ageEdu = ggarrange(p1, p2)
pFEV1_ageEdu

p1
p2



```


## sensitivity analysis with inclusion of BW

## using NN method 

```{r, message=FALSE, warning=FALSE}


library(haven)
PROGRESS_ACCESS_Lung_function_PM_Temp_ <- read_sav("PROGRESS ACCESS PM temp month avgs 020722 small_3.sav")

main<-subset(PROGRESS_ACCESS_Lung_function_PM_Temp_, caliper3>=1)

attach(main)

```


## Create penalized crossbasis 

```{r}

pmweek<-as.matrix (main[,5242:5315])

                         
pm_exp_cb <- crossbasis(pmweek, 
                          lag = c(1, 74), 
                          argvar = list(fun = 'lin'),
                          arglag = list(fun = 'cr'))




tempweek<-as.matrix (main[,5328:5401])

                         
temp_exp_cb <- crossbasis(tempweek, 
                          lag = c(1, 74), 
                          argvar = list(fun = 'lin'),
                          arglag = list(fun = 'cr'))
                          
pen1 <- cbPen(pm_exp_cb )
pen2 <- cbPen(temp_exp_cb)


```





## Method=REML

## FEV model 


```{r, echo=TRUE, warning=FALSE}





mod  <- gam(z_pre_fev1 ~pm_exp_cb + + temp_exp_cb   +as.factor (ets_postnatal) +mom_age +as.factor (educ_dich) + Fenton_Z_score00   ,paraPen = list(pm_exp_cb  = pen1,  
                          temp_exp_cb  = pen2), 
           method = 'REML') # Estimation technique
summary (mod)

# Summarize results
pre_exp_cr <- crossreduce(pm_exp_cb , # DLM crossbasis used in model
                          mod, # Model results
                          type = "var", # Reduction across exposure values
                          value = 5, # Effect relative to exposure value = 1
                          ci.level = 0.95) # Confidence level
post_exp_cr <- crossreduce(temp_exp_cb , mod, type = "var", value = 5)
pre_dlm_results <- data.frame("fit" = pre_exp_cr$fit, 
                              "low" = pre_exp_cr$low,
                              "high" = pre_exp_cr$high)
post_dlm_results <- data.frame("fit" = post_exp_cr$fit, 
                              "low" = post_exp_cr$low,
                              "high" = post_exp_cr$high)



# name rows with '*' to specify critical windows
row.names(pre_dlm_results) <- ifelse(pre_dlm_results$low > 0 | pre_dlm_results$high < 0, paste0("*", row.names(pre_dlm_results)), row.names(pre_dlm_results))
row.names(post_dlm_results) <- ifelse(post_dlm_results$low > 0 | post_dlm_results$high < 0, paste0("*", row.names(post_dlm_results)), row.names(post_dlm_results))
```



```{r, warning=FALSE,message=FALSE}
round(pre_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```

```{r, warning=FALSE,message=FALSE}
round(post_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```




```{r, warning=FALSE,message=FALSE}
pre_dlm_results$lag = 1:74
p1 <- ggplot(pre_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEV1 z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 
 

post_dlm_results$lag = 1:74
p2 <- ggplot(post_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEV1 z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 

pFEV1_ageEdu = ggarrange(p1, p2)
pFEV1_ageEdu

p1

p2



```





## Method=REML

## FVC model 


```{r, echo=TRUE, warning=FALSE}





mod  <- gam(z_pre_fvc ~pm_exp_cb + + temp_exp_cb   +as.factor (ets_postnatal) +mom_age +as.factor (educ_dich) + Fenton_Z_score00  ,paraPen = list(pm_exp_cb  = pen1,  
                          temp_exp_cb  = pen2), 
           method = 'REML') # Estimation technique

summary (mod)


# Summarize results
pre_exp_cr <- crossreduce(pm_exp_cb , # DLM crossbasis used in model
                          mod, # Model results
                          type = "var", # Reduction across exposure values
                          value = 5, # Effect relative to exposure value = 1
                          ci.level = 0.95) # Confidence level
post_exp_cr <- crossreduce(temp_exp_cb , mod, type = "var", value = 5)
pre_dlm_results <- data.frame("fit" = pre_exp_cr$fit, 
                              "low" = pre_exp_cr$low,
                              "high" = pre_exp_cr$high)
post_dlm_results <- data.frame("fit" = post_exp_cr$fit, 
                              "low" = post_exp_cr$low,
                              "high" = post_exp_cr$high)
# name rows with '*' to specify critical windows
row.names(pre_dlm_results) <- ifelse(pre_dlm_results$low > 0 | pre_dlm_results$high < 0, paste0("*", row.names(pre_dlm_results)), row.names(pre_dlm_results))
row.names(post_dlm_results) <- ifelse(post_dlm_results$low > 0 | post_dlm_results$high < 0, paste0("*", row.names(post_dlm_results)), row.names(post_dlm_results))
```


```{r, warning=FALSE,message=FALSE}
round(pre_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```

```{r, warning=FALSE,message=FALSE}
round(post_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```




```{r, warning=FALSE,message=FALSE}
pre_dlm_results$lag = 1:74
p1 <- ggplot(pre_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FVC z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 
 

post_dlm_results$lag = 1:74
p2 <- ggplot(post_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FVC z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 

pFEV1_ageEdu = ggarrange(p1, p2)
pFEV1_ageEdu

p1
p2


```




## Method=REML

## FEV1/FVC model 


```{r, echo=TRUE, warning=FALSE}





mod  <- gam(z_pre_fev1_fvc2 ~pm_exp_cb + + temp_exp_cb   +as.factor (ets_postnatal) +mom_age +as.factor (educ_dich) + Fenton_Z_score00  ,paraPen = list(pm_exp_cb  = pen1,  
                          temp_exp_cb  = pen2), 
           method = 'REML') # Estimation technique

summary (mod)


# Summarize results
pre_exp_cr <- crossreduce(pm_exp_cb , # DLM crossbasis used in model
                          mod, # Model results
                          type = "var", # Reduction across exposure values
                          value = 5, # Effect relative to exposure value = 1
                          ci.level = 0.95) # Confidence level
post_exp_cr <- crossreduce(temp_exp_cb , mod, type = "var", value = 5)
pre_dlm_results <- data.frame("fit" = pre_exp_cr$fit, 
                              "low" = pre_exp_cr$low,
                              "high" = pre_exp_cr$high)
post_dlm_results <- data.frame("fit" = post_exp_cr$fit, 
                              "low" = post_exp_cr$low,
                              "high" = post_exp_cr$high)
# name rows with '*' to specify critical windows
row.names(pre_dlm_results) <- ifelse(pre_dlm_results$low > 0 | pre_dlm_results$high < 0, paste0("*", row.names(pre_dlm_results)), row.names(pre_dlm_results))
row.names(post_dlm_results) <- ifelse(post_dlm_results$low > 0 | post_dlm_results$high < 0, paste0("*", row.names(post_dlm_results)), row.names(post_dlm_results))
```


```{r, warning=FALSE,message=FALSE}
round(pre_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```

```{r, warning=FALSE,message=FALSE}
round(post_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```




```{r, warning=FALSE,message=FALSE}
pre_dlm_results$lag = 1:74
p1 <- ggplot(pre_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEV1/FVC z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 
 

post_dlm_results$lag = 1:74
p2 <- ggplot(post_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEV1/FVC z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 

pFEV1_ageEdu = ggarrange(p1, p2)
pFEV1_ageEdu

p1
p2


```





## Method=REML

## FEF model 


```{r, echo=TRUE, warning=FALSE}





mod  <- gam(z_pre_fef ~pm_exp_cb + + temp_exp_cb   +as.factor (ets_postnatal) +mom_age +as.factor (educ_dich) + Fenton_Z_score00   ,paraPen = list(pm_exp_cb  = pen1,  
                          temp_exp_cb  = pen2), 
           method = 'REML') # Estimation technique

summary (mod)


# Summarize results
pre_exp_cr <- crossreduce(pm_exp_cb , # DLM crossbasis used in model
                          mod, # Model results
                          type = "var", # Reduction across exposure values
                          value = 5, # Effect relative to exposure value = 1
                          ci.level = 0.95) # Confidence level
post_exp_cr <- crossreduce(temp_exp_cb , mod, type = "var", value = 5)
pre_dlm_results <- data.frame("fit" = pre_exp_cr$fit, 
                              "low" = pre_exp_cr$low,
                              "high" = pre_exp_cr$high)
post_dlm_results <- data.frame("fit" = post_exp_cr$fit, 
                              "low" = post_exp_cr$low,
                              "high" = post_exp_cr$high)
# name rows with '*' to specify critical windows
row.names(pre_dlm_results) <- ifelse(pre_dlm_results$low > 0 | pre_dlm_results$high < 0, paste0("*", row.names(pre_dlm_results)), row.names(pre_dlm_results))
row.names(post_dlm_results) <- ifelse(post_dlm_results$low > 0 | post_dlm_results$high < 0, paste0("*", row.names(post_dlm_results)), row.names(post_dlm_results))
```


```{r, warning=FALSE,message=FALSE}
round(pre_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```

```{r, warning=FALSE,message=FALSE}
round(post_dlm_results, 3) %>% kable(format = "html", escape = F) %>%
  kable_styling(font_size = 10) %>% kableExtra::scroll_box(width = "100%", height = "400px")
```




```{r, warning=FALSE,message=FALSE}
pre_dlm_results$lag = 1:74
p1 <- ggplot(pre_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEF z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 
 

post_dlm_results$lag = 1:74
p2 <- ggplot(post_dlm_results, aes(x = lag, y = fit))  + 
  geom_ribbon(aes(ymin = low, ymax = high), fill = "grey70") + 
  geom_line(color="black", size=1) + geom_vline(xintercept = c(2), linetype="dotted", 
                color = "black", size=1)+
  geom_hline(yintercept = 0, color = "black")+  scale_x_continuous(breaks=c(2, 14, 26, 38, 50, 62, 74),label = c("Birth","1", "2", "3",  "4", "5", "6"))+

  theme_classic()+ xlab("Years") + ylab("FEF z-score")+
  theme(axis.text=element_text(size=15,  face="bold"))+ theme(axis.title.x=element_text(size=15,  face="bold")) + theme(axis.title.y=element_text(size=15,  face="bold")) 

pFEV1_ageEdu = ggarrange(p1, p2)
pFEV1_ageEdu

p1
p2



```

